---
output: pdf_document
editor_options: 
  chunk_output_type: inline
---
\mainmatter

# Introduction {#intro}

<!-- **Tips:** -->

<!-- You can label chapter and section titles using `{#label}` after them, e.g., we can reference Chapter \@ref(intro). If you do not manually label them, there will be automatic labels anyway, e.g., Chapter \@ref(methods). -->

<!-- Figures and tables with captions will be placed in `figure` and `table` environments, respectively. -->

<!-- Reference a figure by its code chunk label with the `fig:` prefix, e.g., see Figure \@ref(fig:nice-fig). Similarly, you can reference tables generated from `knitr::kable()`, e.g., see Table \@ref(tab:nice-tab). -->

<!-- You can write citations, too. For example, we are using the **bookdown** package [@R-bookdown] in this sample book, which was built on top of R Markdown and **knitr** [@xie2015]. -->

**Content:**

- Describe Goals of thesis and relation to previous work (Determinants of the price dividend ratio)
    
    - Extend estimation of latent state (mean price dividend ratio) by including possible economic determinants
    
- Describe the purpose of the MCMC Extension (benefits and possible disadvantages)

    - Advantage because of short time series (prior information from previous study)???
    
    - verification with simulation study (linear and nonlinear case)

```{r, import_package_functions, include=FALSE}
import::from(magrittr, "%>%")
```


```{r, read_data, include=FALSE, echo=FALSE}
# read in data
dataset <- readODS::read_ods(here::here("data", "final_dataset.ods"))
relevant_columns <- c(
    "Date",
    "Real Consumption (per Capita)",
    "Consumption (per Capita)",
    "USPOPTO",
    "USPOP24Y",
    "USPOP29Y",
    "USPOP44Y",
    "USPOP49Y",
    "S&P",
    "Dividend",
    "Earnings",
    "CPI",
    "rf",
    "fund_rate",
    "real_m1"
)
# filter relevant columns
dataset <- dataset[, relevant_columns]

# rename columns
renamed_columns <- c(
    "date",
    "real_consumption",
    "consumption",
    "us_pop_total",
    "us_pop_20_24",
    "us_pop_25_29",
    "us_pop_40_44",
    "us_pop_45_49",
    "price",
    "dividend",
    "earnings",
    "cpi",
    "rf",
    "fund_rate",
    "real_m1"
)
colnames(dataset) <- renamed_columns
row.names(dataset) <- dataset$date
dataset <- dataset[, colnames(dataset)[2:length(colnames(dataset))]]

# transform to xts object
dataset <- xts::xts(
    dataset, order.by = as.Date(row.names(dataset), "%Y-%m-%d")
)
dataset <- dataset["1949-12-31/2019-12-31"]

# linear interpolation for population data
pop_columns <- c(
    "us_pop_total",
    "us_pop_20_24",
    "us_pop_25_29",
    "us_pop_40_44",
    "us_pop_45_49"
)
dataset[
    -dataset[xts:::endof(dataset, "years"), which.i = TRUE],
    pop_columns
] <- NA
storage.mode(dataset) <- "numeric"

dataset[, pop_columns] <- zoo::na.approx(dataset[, pop_columns])
dataset <- dataset[complete.cases(dataset), ]
# dataset <- dataset[xts:::endof(dataset, "years")]

head(dataset)
```

```{r, consumption_volatility, include=FALSE, eval=TRUE}
con <- dataset$real_consumption
quar_c_growth <- diff(con) / lag(con, k = 1)
quar_c_growth <- quar_c_growth[!is.na(quar_c_growth)]
quar_c_growth <- quar_c_growth - mean(quar_c_growth)
con_vol <- log(zoo::rollapply(abs(quar_c_growth), width = 20, FUN = sum, 
                              align = "right"))
```

```{r, middle_aged_young_ratio, include=FALSE, eval=TRUE}
middle_aged <- dataset$us_pop_40_44 + dataset$us_pop_45_49
young <- dataset$us_pop_20_24 + dataset$us_pop_25_29
middle_aged_young_ratio <- middle_aged / young
```

## Model

**Measurement:**

$$\eta_{t} = \frac{k_{t}}{1-\rho_{t}}+\sum_{i\;=\;1}^{\infty}\rho^{i-1}_{t}\tilde{E}[\Delta d^{e}_{t+i}-r_{t+i}^{e}] + \epsilon_{t}$$

$$\epsilon_{t} \sim N(0, \sigma_{\epsilon}^2)$$

$$\rho_{t} = \frac{1}{1+exp(-\tilde{\eta_{t}})}$$

$$k_{t} = -ln(\rho_{t})-(1-\rho_{t})ln\bigg(\frac{1}{\rho_{t}} - 1\bigg)$$

**State:**

$$\tilde{\eta}_{t} = \tilde{\eta}_{t-1} + u_{t}$$

**Model with covariates:**

$$\tilde{\eta}_{t} = \beta_{0} + \beta_{1}\tilde{\eta}_{t-1} + \beta_{2}cr_{t-1} + \beta_{3}my_{t-1} + \beta_{4}fr_{t-1} + \beta_{5}ms_{t-1} + u_{t}$$

$$u_{t} \sim N(0, \sigma_{u}^2)$$

\newpage

## Estimation

**Data:**

- *lpd*: log price dividend ratio
- *d_e*: excess dividends
- *r_e*: excess returns
- *infl*: inflation
- *cr*: consumption risk
- *my*: middle aged to young ratio
- *fr*: effective federal funds rate
- *ms*: log real money supply m1

```{r, data_for_estimation, echo=FALSE}
sub_dataset <- dataset

r_e <- (
  log((sub_dataset$price + sub_dataset$dividend) / 
        stats::lag(sub_dataset$price, k = 1))
  - log(1 + (sub_dataset$rf / 100))
)
d_e <- (
  log(sub_dataset$dividend / stats::lag(sub_dataset$dividend, k = 1))
  - log(1 + (sub_dataset$rf / 100))
)
con_vol <- con_vol
my <- middle_aged_young_ratio
fr <- log(1 + (sub_dataset$fund_rate/100))
ms <- log(sub_dataset$real_m1)
inflation <- (diff(sub_dataset$cpi) / stats::lag(sub_dataset$cpi, k = 1))
lpd <- log((sub_dataset$price / sub_dataset$dividend))

var_data <- cbind(lpd, d_e, r_e, inflation, con_vol, my, fr, ms)
var_data <- var_data[complete.cases(var_data),]

starting_point <- 41
params <- matrix(nrow = nrow(var_data) - starting_point + 1, ncol = 20)
for (i in starting_point:nrow(var_data)) {
    residuals_model <- vector("numeric", length = 11)
    for (w in 30:40) {
        var_model <- vars::VAR(var_data[(i - w + 1):i, 1:4], p = 1,
                               type = "const")
        residuals_j <- tail(residuals(var_model), 10)[, c(2, 3)]
        residuals_model[w - 29] <- sqrt(
            mean((residuals_j[, 1] - residuals_j[, 2])^2)
        )
    }
    best_model <- which.min(residuals_model)
    model <- vars::VAR(var_data[(i - best_model - 29):i, 1:4], type = "const")
    params[i - starting_point + 1, ] <- as.vector(t(vars::Bcoef(model)))
}

var_data <- cbind(var_data[starting_point:nrow(var_data), ], params)
y <- var_data[complete.cases(var_data),]
y <- cbind(1:nrow(y), y)
colnames(y) <- c("time", "lpd", "d_e", "r_e", "inflation", "cr", "mys", "fr", 
                 "ms", sapply(1:20, function(x){sprintf("param_%i", x)}))
dates <- zoo::index(y)
y <- as.data.frame(y, row.names = NULL)

head(y, 10)
```

```{r}
test <- pomp::pomp(
    data = y[, c(1, 2)],
    times = "time", t0 = 1,
    params = ,
    rinit = function(e_lpd_0, ...) {
        return(c(e_lpd = e_lpd_0))
    },
    rprocess = pomp::discrete_time(pomp::Csnippet(
      "e_lpd = e_lpd + rnorm(0, sigma_u);"
    ), delta.t = 1),
    dmeasure = dm,
    statenames = "e_lpd",
    partrans = pomp::parameter_trans(log = c("sigma_u", "sigma_e")),
    paramnames = c("e_lpd_0", "sigma_u", "sigma_e"),
    covar = pomp::covariate_table(y[, -c(2, 6, 7, 8, 9)], times = "time"),
    covarnames = colnames(y[, -c(1, 2, 6, 7, 8, 9)])
)

theta <- c(sigma_u = 0.05, e_lpd_0 = 3.5, sigma_e = 0.05)

pf <- pomp::pfilter(test, params = theta, Np = 250, 
                    pred.mean = TRUE, pred.var = TRUE, 
                    filter.mean = T, filter.traj = TRUE, 
                    save.states = TRUE, verbose = TRUE)
pomp::logLik(pf)

m <- pomp::mif2(
  test, Nmif = 50, Np = 250, params = theta, cooling.fraction = 0.7,
  rw.sd = pomp::rw.sd(e_lpd_0 = ivp(3.5), sigma_e = 0.02, sigma_u = 0.02),
  verbose = TRUE
)

nm <- optim(
  par = c(sigma_u = log(0.05), e_lpd_0 = 3.5, sigma_e = log(0.05)),
  fn = function(par, ...){
    par[c(1, 3)] <- exp(par[c(1, 3)])
    print(par)
    return(-logLik(pfilter(test, params = par, Np = 250)))
  },
  test = test
)

theta1 <- nm$par
theta1[c(1, 3)] <- exp(theta1[c(1, 3)])
pf <- pfilter(test, params = theta1, Np = 250, 
              pred.mean = TRUE, pred.var = TRUE, 
              filter.mean = T, filter.traj = TRUE, 
              save.states = TRUE, verbose = TRUE)
logLik(pf)
plot(pf)
```
