---
output: html_document
editor_options: 
  chunk_output_type: console
---
# (APPENDIX) Appendix {-}

# Appendix: Figures {#figures}

```{r, sim-4-appendix, eval=TRUE, echo=FALSE}
# get simulation result
source(here::here("R", "simulation_study", "simulation.R"))

# Plot trace noninformative
sim_4 <- simulation_result[[4]]

library(ggplot2)
library(gridExtra)
```

```{r, sim-4-appendix-trace-plot-setup-1, eval=TRUE, echo=FALSE, fig.align='center', fig.cap="Exemplary simulation: Trace plot setup 1", fig.width=6, fig.height=6.5, warning=FALSE}
traces <- sim_4$pmmh_noninformative$traces

# Plot theme:
p_theme <- theme(
  legend.justification = c(0, 1), legend.position = c(0, 1), 
  legend.title = element_blank(), 
  legend.background = element_rect(fill = "transparent"),
  plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
  plot.title = element_text(hjust = 0.5, size = 8),
  axis.title.x = element_text(size = 8),
  axis.title.y = element_text(size = 8, angle = 0, vjust = 0.5)
)

# Beta 0
df <- data.frame(
  time = 1:nrow(traces), variable = "beta_0", beta_0 = traces[, 3]
)
p1 <- ggplot(data = df, aes(x = time, y = beta_0)) +
  geom_line(aes(colour = variable)) + theme_light() +
  geom_hline(
    aes(yintercept = true_params_simulation["beta_0"], color = "True"),
    show.legend = TRUE, linetype = "dashed"
  ) +
  scale_color_manual(values = c("black", "red"), labels = c("Trace", "True")) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\beta_0$}"),
       x = "m", y = latex2exp::TeX("$\\beta_0$")) +
  p_theme

# Beta 1
df <- data.frame(
  time = 1:nrow(traces), beta_1 = traces[, 4], variable = "Trace"
)
p2 <- ggplot(data = df, aes(x = time, y = beta_1)) +
  geom_line(aes(colour = variable)) + theme_light() +
  geom_hline(
    aes(yintercept = true_params_simulation["beta_1"], color = "True"),
    show.legend = TRUE, linetype = "dashed"
  ) +
  scale_color_manual(values = c("black", "red"), labels = c("Trace", "True")) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\beta_1$}"),
       x = "m", y = latex2exp::TeX("$\\beta_1$")) +
  p_theme

# Beta 2
df <- data.frame(
  time = 1:nrow(traces), beta_2 = traces[, 5], variable = "Trace"
)
p3 <- ggplot(data = df, aes(x = time, y = beta_2)) +
  geom_line(aes(colour = variable)) + theme_light() +
  geom_hline(
    aes(yintercept = true_params_simulation["beta_2"], color = "True"),
    show.legend = TRUE, linetype = "dashed"
  ) +
  scale_color_manual(values = c("black", "red"), labels = c("Trace", "True")) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\beta_2$}"),
       x = "m", y = latex2exp::TeX("$\\beta_2$")) +
  p_theme

# Sigma x
df <- data.frame(
  time = 1:nrow(traces), sigma_x = exp(traces[, 6]), variable = "Trace"
)
p4 <- ggplot(data = df, aes(x = time, y = sigma_x)) +
  geom_line(aes(colour = variable)) + theme_light() +
  geom_hline(
    aes(yintercept = exp(true_params_simulation["sigma_x"]), color = "True"),
    show.legend = TRUE, linetype = "dashed"
  ) +
  scale_color_manual(values = c("black", "red"), labels = c("Trace", "True")) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\sigma_u$}"),
       x = "m", y = latex2exp::TeX("$\\sigma_u$")) +
  p_theme

# Sigma y
df <- data.frame(
  time = 1:nrow(traces), sigma_y = exp(traces[, 7]), variable = "Trace"
)
p5 <- ggplot(data = df, aes(x = time, y = sigma_y)) +
  geom_line(aes(colour = variable)) + theme_light() +
  geom_hline(
    aes(yintercept = exp(true_params_simulation["sigma_y"]), color = "True"),
    show.legend = TRUE, linetype = "dashed"
  ) +
  scale_color_manual(values = c("black", "red"), labels = c("Trace", "True")) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\sigma_\\epsilon$}"),
       x = "m", y = latex2exp::TeX("$\\sigma_\\epsilon$")) +
  p_theme

# Phi
df <- data.frame(
  time = 1:nrow(traces), phi = tanh(traces[, 8]), variable = "Trace"
)
p6 <- ggplot(data = df, aes(x = time, y = phi)) +
  geom_line(aes(colour = variable)) + theme_light() +
  geom_hline(
    aes(yintercept = tanh(true_params_simulation["phi"]), color = "True"),
    show.legend = TRUE, linetype = "dashed"
  ) +
  scale_color_manual(values = c("black", "red"), labels = c("Trace", "True")) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\phi$}"),
       x = "m", y = latex2exp::TeX("$\\phi$")) +
  p_theme

# X 0
df <- data.frame(
  time = 1:nrow(traces), x_0 = traces[, 9], variable = "Trace"
)
p7 <- ggplot(data = df, aes(x = time, y = x_0)) +
  geom_line(aes(colour = variable)) + theme_light() +
  geom_hline(
    aes(yintercept = true_params_simulation["x_0"], color = "True"),
    show.legend = TRUE, linetype = "dashed"
  ) +
  scale_color_manual(values = c("black", "red"), labels = c("Trace", "True")) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\x_0$}"),
       x = "m", y = latex2exp::TeX("$\\x_0$")) +
  p_theme

get_legend <- function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}
legend <- get_legend(p2)

grid.arrange(
  p1 + theme(legend.position = "none"), 
  p2 + theme(legend.position = "none"), 
  p3 + theme(legend.position = "none"), 
  p4 + theme(legend.position = "none"), 
  p5 + theme(legend.position = "none"), 
  p6 + theme(legend.position = "none"), 
  p7 + theme(legend.position = "none"), 
  ncol = 2, legend = legend
)
```

```{r, sim-4-appendix-marginal-posterior-setup-1, eval=TRUE, echo=FALSE, fig.align='center', fig.cap="Exemplary simulation: Gaussian kernel density estimates of the marginal posterior distributions for setup 1", fig.width=6, fig.height=6.5, warning=FALSE}
traces <- sim_4$pmmh_noninformative$traces[10000:20000, ]

# Plot theme:
p_theme <- theme(
  legend.justification = c(0, 1), legend.position = c(0, 1), 
  legend.title = element_blank(), 
  legend.background = element_rect(fill = "transparent"),
  plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
  plot.title = element_text(hjust = 0.5, size = 8),
  axis.title.x = element_text(size = 8),
  axis.title.y = element_text(size = 8, vjust = 0.5),
  legend.key = element_rect(fill = NA)
)

# Beta 0
df <- data.frame(time = 1:nrow(traces), beta_0 = traces[, 3])
p1 <- ggplot(data = df, aes(x = beta_0)) +
  geom_density(aes(x = beta_0, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = true_params_simulation["beta_0"], colour = "True",
        linetype = "True"), show.legend = FALSE
  ) +
  geom_vline(
    aes(xintercept = mean(traces[, 3]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", "True" = "red",
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, 
               "Prior" = 1, "True" = 3), name = ""
  ) + 
  labs(title = latex2exp::TeX("Marginal Posterior: \\textbf{$\\beta_0$}"),
       x = latex2exp::TeX("$\\beta_0$"))

# Beta 1
df <- data.frame(time = 1:nrow(traces), beta_1 = traces[, 4])
p2 <- ggplot(data = df, aes(x = beta_1)) +
  geom_density(aes(x = beta_1, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = true_params_simulation["beta_1"], colour = "True",
        linetype = "True"), show.legend = FALSE
  ) +
  geom_vline(
    aes(xintercept = mean(traces[, 4]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", "True" = "red",
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, 
               "Prior" = 1, "True" = 3), name = ""
  ) + 
  labs(title = latex2exp::TeX("Marginal Posterior: \\textbf{$\\beta_1$}"),
       x = latex2exp::TeX("$\\beta_1$"))

# Beta 2
df <- data.frame(time = 1:nrow(traces), beta_2 = traces[, 5])
p3 <- ggplot(data = df, aes(x = beta_2)) +
  geom_density(aes(x = beta_2, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = true_params_simulation["beta_2"], colour = "True",
        linetype = "True"), show.legend = FALSE
  ) +
  geom_vline(
    aes(xintercept = mean(traces[, 5]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", "True" = "red",
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, 
               "Prior" = 1, "True" = 3), name = ""
  ) + 
  labs(title = latex2exp::TeX("Marginal Posterior: \\textbf{$\\beta_2$}"),
       x = latex2exp::TeX("$\\beta_2$"))

# Sigma x
df <- data.frame(time = 1:nrow(traces), sigma_x = exp(traces[, 6]))
p4 <- ggplot(data = df, aes(x = sigma_x)) +
  geom_density(aes(x = sigma_x, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = exp(true_params_simulation["sigma_x"]), colour = "True",
        linetype = "True"), show.legend = FALSE
  ) +
  geom_vline(
    aes(xintercept = mean(exp(traces[, 6])), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", "True" = "red",
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, 
               "Prior" = 1, "True" = 3), name = ""
  ) + 
  labs(title = latex2exp::TeX("Marginal Posterior: \\textbf{$\\sigma_u$}"),
       x = latex2exp::TeX("$\\sigma_u$"))

# Sigma y
df <- data.frame(time = 1:nrow(traces), sigma_y = exp(traces[, 7]))
p5 <- ggplot(data = df, aes(x = sigma_y)) +
  geom_density(aes(x = sigma_y, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = exp(true_params_simulation["sigma_y"]), colour = "True",
        linetype = "True"), show.legend = FALSE
  ) +
  geom_vline(
    aes(xintercept = mean(exp(traces[, 7])), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", "True" = "red",
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, 
               "Prior" = 1, "True" = 3), name = ""
  ) + 
  labs(title = latex2exp::TeX("Marginal Posterior: \\textbf{$\\sigma_\\epsilon$}"),
       x = latex2exp::TeX("$\\sigma_\\epsilon$"))

# Phi
df <- data.frame(time = 1:nrow(traces), phi = tanh(traces[, 8]))
p6 <- ggplot(data = df, aes(x = phi)) +
  geom_density(aes(x = phi, colour = "Posterior", linetype = "Posterior"),
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = tanh(true_params_simulation["phi"]), colour = "True",
        linetype = "True"), show.legend = FALSE
  ) +
  geom_vline(
    aes(xintercept = mean(tanh(traces[, 8])), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", "True" = "red",
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, 
               "Prior" = 1, "True" = 3), name = ""
  ) + 
  labs(title = latex2exp::TeX("Marginal Posterior: \\textbf{$\\phi$}"),
       x = latex2exp::TeX("$\\phi$"))

# X 0
df <- data.frame(time = 1:nrow(traces), x_0 = traces[, 9])
p7 <- ggplot(data = df, aes(x = x_0)) +
  geom_density(aes(x = x_0, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = true_params_simulation["x_0"], colour = "True",
        linetype = "True"), show.legend = FALSE
  ) +
  geom_vline(
    aes(xintercept = mean(traces[, 9]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", "True" = "red",
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, 
               "Prior" = 1, "True" = 3), name = ""
  ) + 
  labs(title = latex2exp::TeX("Marginal Posterior: \\textbf{$\\x_0$}"),
       x = latex2exp::TeX("$\\x_0$"))

get_legend <- function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}
legend <- get_legend(p2)

grid.arrange(
  p1 + theme(legend.position = "none"), 
  p2 + theme(legend.position = "none"), 
  p3 + theme(legend.position = "none"), 
  p4 + theme(legend.position = "none"), 
  p5 + theme(legend.position = "none"), 
  p6 + theme(legend.position = "none"), 
  p7 + theme(legend.position = "none"), 
  ncol = 2, legend = legend
)
```

```{r, sim-4-appendix-trace-plot-setup-2, eval=TRUE, echo=FALSE, fig.align='center', fig.cap="Exemplary simulation: Trace plot setup 2", fig.width=6, fig.height=6.5, warning=FALSE}
traces <- sim_4$pmmh_true_informative$traces

# Plot theme:
p_theme <- theme(
  legend.justification = c(0, 1), legend.position = c(0, 1), 
  legend.title = element_blank(), 
  legend.background = element_rect(fill = "transparent"),
  plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
  plot.title = element_text(hjust = 0.5, size = 8),
  axis.title.x = element_text(size = 8),
  axis.title.y = element_text(size = 8, angle = 0, vjust = 0.5)
)

# Beta 0
df <- data.frame(
  time = 1:nrow(traces), variable = "beta_0", beta_0 = traces[, 3]
)
p1 <- ggplot(data = df, aes(x = time, y = beta_0)) +
  geom_line(aes(colour = variable)) + theme_light() +
  geom_hline(
    aes(yintercept = true_params_simulation["beta_0"], color = "True"),
    show.legend = TRUE, linetype = "dashed"
  ) +
  scale_color_manual(values = c("black", "red"), labels = c("Trace", "True")) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\beta_0$}"),
       x = "m", y = latex2exp::TeX("$\\beta_0$")) +
  p_theme

# Beta 1
df <- data.frame(
  time = 1:nrow(traces), beta_1 = traces[, 4], variable = "Trace"
)
p2 <- ggplot(data = df, aes(x = time, y = beta_1)) +
  geom_line(aes(colour = variable)) + theme_light() +
  geom_hline(
    aes(yintercept = true_params_simulation["beta_1"], color = "True"),
    show.legend = TRUE, linetype = "dashed"
  ) +
  scale_color_manual(values = c("black", "red"), labels = c("Trace", "True")) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\beta_1$}"),
       x = "m", y = latex2exp::TeX("$\\beta_1$")) +
  p_theme

# Beta 2
df <- data.frame(
  time = 1:nrow(traces), beta_2 = traces[, 5], variable = "Trace"
)
p3 <- ggplot(data = df, aes(x = time, y = beta_2)) +
  geom_line(aes(colour = variable)) + theme_light() +
  geom_hline(
    aes(yintercept = true_params_simulation["beta_2"], color = "True"),
    show.legend = TRUE, linetype = "dashed"
  ) +
  scale_color_manual(values = c("black", "red"), labels = c("Trace", "True")) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\beta_2$}"),
       x = "m", y = latex2exp::TeX("$\\beta_2$")) + 
  p_theme

# Sigma x
df <- data.frame(
  time = 1:nrow(traces), sigma_x = exp(traces[, 6]), variable = "Trace"
)
p4 <- ggplot(data = df, aes(x = time, y = sigma_x)) +
  geom_line(aes(colour = variable)) + theme_light() +
  geom_hline(
    aes(yintercept = exp(true_params_simulation["sigma_x"]), color = "True"),
    show.legend = TRUE, linetype = "dashed"
  ) +
  scale_color_manual(values = c("black", "red"), labels = c("Trace", "True")) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\sigma_u$}"),
       x = "m", y = latex2exp::TeX("$\\sigma_u$")) +
  p_theme

# Sigma y
df <- data.frame(
  time = 1:nrow(traces), sigma_y = exp(traces[, 7]), variable = "Trace"
)
p5 <- ggplot(data = df, aes(x = time, y = sigma_y)) +
  geom_line(aes(colour = variable)) + theme_light() +
  geom_hline(
    aes(yintercept = exp(true_params_simulation["sigma_y"]), color = "True"),
    show.legend = TRUE, linetype = "dashed"
  ) +
  scale_color_manual(values = c("black", "red"), labels = c("Trace", "True")) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\sigma_\\epsilon$}"),
       x = "m", y = latex2exp::TeX("$\\sigma_\\epsilon$")) +
  p_theme

# Phi
df <- data.frame(
  time = 1:nrow(traces), phi = tanh(traces[, 8]), variable = "Trace"
)
p6 <- ggplot(data = df, aes(x = time, y = phi)) +
  geom_line(aes(colour = variable)) + theme_light() +
  geom_hline(
    aes(yintercept = tanh(true_params_simulation["phi"]), color = "True"),
    show.legend = TRUE, linetype = "dashed"
  ) +
  scale_color_manual(values = c("black", "red"), labels = c("Trace", "True")) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\phi$}"),
       x = "m", y = latex2exp::TeX("$\\phi$")) +
  p_theme

# X 0
df <- data.frame(
  time = 1:nrow(traces), x_0 = traces[, 9], variable = "Trace"
)
p7 <- ggplot(data = df, aes(x = time, y = x_0)) +
  geom_line(aes(colour = variable)) + theme_light() +
  geom_hline(
    aes(yintercept = true_params_simulation["x_0"], color = "True"),
    show.legend = TRUE, linetype = "dashed"
  ) +
  scale_color_manual(values = c("black", "red"), labels = c("Trace", "True")) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\x_0$}"),
       x = "m", y = latex2exp::TeX("$\\x_0$")) +
  p_theme

get_legend <- function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}
legend <- get_legend(p2)

grid.arrange(
  p1 + theme(legend.position = "none"), 
  p2 + theme(legend.position = "none"), 
  p3 + theme(legend.position = "none"), 
  p4 + theme(legend.position = "none"), 
  p5 + theme(legend.position = "none"), 
  p6 + theme(legend.position = "none"), 
  p7 + theme(legend.position = "none"), 
  ncol = 2, legend = legend
)
```

```{r, sim-4-appendix-marginal-posterior-setup-2, eval=TRUE, echo=FALSE, fig.align='center', fig.cap="Exemplary simulation: Gaussian kernel density estimates of the marginal posterior distributions for setup 2", fig.width=6, fig.height=6.5, warning=FALSE}
traces <- sim_4$pmmh_true_informative$traces[10000:20000, ]

# Plot theme:
p_theme <- theme(
  legend.justification = c(0, 1), legend.position = c(0, 1), 
  legend.title = element_blank(), 
  legend.background = element_rect(fill = "transparent"),
  plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
  plot.title = element_text(hjust = 0.5, size = 8),
  axis.title.x = element_text(size = 8),
  axis.title.y = element_text(size = 8, vjust = 0.5),
  legend.key = element_rect(fill = NA)
)

# Beta 0
df <- data.frame(time = 1:nrow(traces), beta_0 = traces[, 3])
p1 <- ggplot(data = df, aes(x = beta_0)) +
  geom_density(aes(x = beta_0, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = true_params_simulation["beta_0"], colour = "True",
        linetype = "True"), show.legend = FALSE
  ) +
  geom_vline(
    aes(xintercept = mean(traces[, 3]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = dnorm, aes(colour = "Prior", linetype = "Prior"),
    args = list(mean = 0.4, sd = 0.1)
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", "True" = "red",
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, 
               "Prior" = 1, "True" = 3), name = ""
  ) + 
  labs(title = latex2exp::TeX("Marginal Posterior: \\textbf{$\\beta_0$}"),
       x = latex2exp::TeX("$\\beta_0$"))

# Beta 1
df <- data.frame(time = 1:nrow(traces), beta_1 = traces[, 4])
p2 <- ggplot(data = df, aes(x = beta_1)) +
  geom_density(aes(x = beta_1, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = true_params_simulation["beta_1"], colour = "True",
        linetype = "True"), show.legend = FALSE
  ) +
  geom_vline(
    aes(xintercept = mean(traces[, 4]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = dnorm, aes(colour = "Prior", linetype = "Prior"),
    args = list(mean = 0.5, sd = 0.2)
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", "True" = "red",
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, 
               "Prior" = 1, "True" = 3), name = ""
  ) + 
  labs(title = latex2exp::TeX("Marginal Posterior: \\textbf{$\\beta_1$}"),
       x = latex2exp::TeX("$\\beta_1$"))

# Beta 2
df <- data.frame(time = 1:nrow(traces), beta_2 = traces[, 5])
p3 <- ggplot(data = df, aes(x = beta_2)) +
  geom_density(aes(x = beta_2, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = true_params_simulation["beta_2"], colour = "True",
        linetype = "True"), show.legend = FALSE
  ) +
  geom_vline(
    aes(xintercept = mean(traces[, 5]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = dnorm, aes(colour = "Prior", linetype = "Prior"),
    args = list(mean = -0.3, sd = 0.2)
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", "True" = "red",
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, 
               "Prior" = 1, "True" = 3), name = ""
  ) + 
  labs(title = latex2exp::TeX("Marginal Posterior: \\textbf{$\\beta_2$}"),
       x = latex2exp::TeX("$\\beta_2$"))

# Sigma x
df <- data.frame(time = 1:nrow(traces), sigma_x = exp(traces[, 6]))
p4 <- ggplot(data = df, aes(x = sigma_x)) +
  geom_density(aes(x = sigma_x, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = exp(true_params_simulation["sigma_x"]), colour = "True",
        linetype = "True"), show.legend = FALSE
  ) +
  geom_vline(
    aes(xintercept = mean(exp(traces[, 6])), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = dgamma, aes(colour = "Prior", linetype = "Prior"),
    args = list(shape = 12, scale = 0.005)
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", "True" = "red",
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, 
               "Prior" = 1, "True" = 3), name = ""
  ) + 
  labs(title = latex2exp::TeX("Marginal Posterior: \\textbf{$\\sigma_u$}"),
       x = latex2exp::TeX("$\\sigma_u$"))

# Sigma y
df <- data.frame(time = 1:nrow(traces), sigma_y = exp(traces[, 7]))
p5 <- ggplot(data = df, aes(x = sigma_y)) +
  geom_density(aes(x = sigma_y, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = exp(true_params_simulation["sigma_y"]), colour = "True",
        linetype = "True"), show.legend = FALSE
  ) +
  geom_vline(
    aes(xintercept = mean(exp(traces[, 7])), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = dgamma, aes(colour = "Prior", linetype = "Prior"),
    args = list(shape = 6, scale = 0.005)
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", "True" = "red",
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, 
               "Prior" = 1, "True" = 3), name = ""
  ) + 
  labs(title = latex2exp::TeX("Marginal Posterior: \\textbf{$\\sigma_\\epsilon$}"),
       x = latex2exp::TeX("$\\sigma_\\epsilon$"))

# Phi
df <- data.frame(time = 1:nrow(traces), phi = tanh(traces[, 8]))
p6 <- ggplot(data = df, aes(x = phi)) +
  geom_density(aes(x = phi, colour = "Posterior", linetype = "Posterior"),
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = tanh(true_params_simulation["phi"]), colour = "True",
        linetype = "True"), show.legend = FALSE
  ) +
  geom_vline(
    aes(xintercept = mean(tanh(traces[, 8])), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = truncnorm::dtruncnorm, aes(colour = "Prior", linetype = "Prior"),
    args = list(a = -1, b = 1, mean = 0.8, sd = 0.1)
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", "True" = "red",
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, 
               "Prior" = 1, "True" = 3), name = ""
  ) + 
  labs(title = latex2exp::TeX("Marginal Posterior: \\textbf{$\\phi$}"),
       x = latex2exp::TeX("$\\phi$"))

# X 0
df <- data.frame(time = 1:nrow(traces), x_0 = traces[, 9])
p7 <- ggplot(data = df, aes(x = x_0)) +
  geom_density(aes(x = x_0, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = true_params_simulation["x_0"], colour = "True",
        linetype = "True"), show.legend = FALSE
  ) +
  geom_vline(
    aes(xintercept = mean(traces[, 9]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = dnorm, aes(colour = "Prior", linetype = "Prior"),
    args = list(mean = 3.5, sd = 0.1)
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", "True" = "red",
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, 
               "Prior" = 1, "True" = 3), name = ""
  ) + 
  labs(title = latex2exp::TeX("Marginal Posterior: \\textbf{$\\x_0$}"),
       x = latex2exp::TeX("$\\x_0$"))

get_legend <- function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}
legend <- get_legend(p2)

grid.arrange(
  p1 + theme(legend.position = "none"), 
  p2 + theme(legend.position = "none"), 
  p3 + theme(legend.position = "none"), 
  p4 + theme(legend.position = "none"), 
  p5 + theme(legend.position = "none"), 
  p6 + theme(legend.position = "none"), 
  p7 + theme(legend.position = "none"), 
  ncol = 2, legend = legend
)
```

```{r, sim-4-appendix-trace-plot-setup-3, eval=TRUE, echo=FALSE, fig.align='center', fig.cap="Exemplary simulation: Trace plot setup 3", fig.width=6, fig.height=6.5, warning=FALSE}
traces <- sim_4$pmmh_false_informative$traces

# Plot theme:
p_theme <- theme(
  legend.justification = c(0, 1), legend.position = c(0, 1), 
  legend.title = element_blank(), 
  legend.background = element_rect(fill = "transparent"),
  plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
  plot.title = element_text(hjust = 0.5, size = 8),
  axis.title.x = element_text(size = 8),
  axis.title.y = element_text(size = 8, angle = 0, vjust = 0.5)
)

# Beta 0
df <- data.frame(
  time = 1:nrow(traces), variable = "beta_0", beta_0 = traces[, 3]
)
p1 <- ggplot(data = df, aes(x = time, y = beta_0)) +
  geom_line(aes(colour = variable)) + theme_light() +
  geom_hline(
    aes(yintercept = true_params_simulation["beta_0"], color = "True"),
    show.legend = TRUE, linetype = "dashed"
  ) +
  scale_color_manual(values = c("black", "red"), labels = c("Trace", "True")) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\beta_0$}"),
       x = "m", y = latex2exp::TeX("$\\beta_0$")) +
  p_theme

# Beta 1
df <- data.frame(
  time = 1:nrow(traces), beta_1 = traces[, 4], variable = "Trace"
)
p2 <- ggplot(data = df, aes(x = time, y = beta_1)) +
  geom_line(aes(colour = variable)) + theme_light() +
  geom_hline(
    aes(yintercept = true_params_simulation["beta_1"], color = "True"),
    show.legend = TRUE, linetype = "dashed"
  ) +
  scale_color_manual(values = c("black", "red"), labels = c("Trace", "True")) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\beta_1$}"),
       x = "m", y = latex2exp::TeX("$\\beta_1$")) +
  p_theme

# Beta 2
df <- data.frame(
  time = 1:nrow(traces), beta_2 = traces[, 5], variable = "Trace"
)
p3 <- ggplot(data = df, aes(x = time, y = beta_2)) +
  geom_line(aes(colour = variable)) + theme_light() +
  geom_hline(
    aes(yintercept = true_params_simulation["beta_2"], color = "True"),
    show.legend = TRUE, linetype = "dashed"
  ) +
  scale_color_manual(values = c("black", "red"), labels = c("Trace", "True")) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\beta_2$}"),
       x = "m", y = latex2exp::TeX("$\\beta_2$")) +
  p_theme

# Sigma x
df <- data.frame(
  time = 1:nrow(traces), sigma_x = exp(traces[, 6]), variable = "Trace"
)
p4 <- ggplot(data = df, aes(x = time, y = sigma_x)) +
  geom_line(aes(colour = variable)) + theme_light() +
  geom_hline(
    aes(yintercept = exp(true_params_simulation["sigma_x"]), color = "True"),
    show.legend = TRUE, linetype = "dashed"
  ) +
  scale_color_manual(values = c("black", "red"), labels = c("Trace", "True")) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\sigma_u$}"),
       x = "m", y = latex2exp::TeX("$\\sigma_u$")) +
  p_theme

# Sigma y
df <- data.frame(
  time = 1:nrow(traces), sigma_y = exp(traces[, 7]), variable = "Trace"
)
p5 <- ggplot(data = df, aes(x = time, y = sigma_y)) +
  geom_line(aes(colour = variable)) + theme_light() +
  geom_hline(
    aes(yintercept = exp(true_params_simulation["sigma_y"]), color = "True"),
    show.legend = TRUE, linetype = "dashed"
  ) +
  scale_color_manual(values = c("black", "red"), labels = c("Trace", "True")) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\sigma_\\epsilon$}"),
       x = "m", y = latex2exp::TeX("$\\sigma_\\epsilon$")) +
  p_theme

# Phi
df <- data.frame(
  time = 1:nrow(traces), phi = tanh(traces[, 8]), variable = "Trace"
)
p6 <- ggplot(data = df, aes(x = time, y = phi)) +
  geom_line(aes(colour = variable)) + theme_light() +
  geom_hline(
    aes(yintercept = tanh(true_params_simulation["phi"]), color = "True"),
    show.legend = TRUE, linetype = "dashed"
  ) +
  scale_color_manual(values = c("black", "red"), labels = c("Trace", "True")) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\phi$}"),
       x = "m", y = latex2exp::TeX("$\\phi$")) +
  p_theme

# X 0
df <- data.frame(
  time = 1:nrow(traces), x_0 = traces[, 9], variable = "Trace"
)
p7 <- ggplot(data = df, aes(x = time, y = x_0)) +
  geom_line(aes(colour = variable)) + theme_light() +
  geom_hline(
    aes(yintercept = true_params_simulation["x_0"], color = "True"),
    show.legend = TRUE, linetype = "dashed"
  ) +
  scale_color_manual(values = c("black", "red"), labels = c("Trace", "True")) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\x_0$}"),
       x = "m", y = latex2exp::TeX("$\\x_0$")) +
  p_theme

get_legend <- function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}
legend <- get_legend(p2)

grid.arrange(
  p1 + theme(legend.position = "none"), 
  p2 + theme(legend.position = "none"), 
  p3 + theme(legend.position = "none"), 
  p4 + theme(legend.position = "none"), 
  p5 + theme(legend.position = "none"), 
  p6 + theme(legend.position = "none"), 
  p7 + theme(legend.position = "none"), 
  ncol = 2, legend = legend
)
```

```{r, sim-4-appendix-marginal-posterior-setup-3, eval=TRUE, echo=FALSE, fig.align='center', fig.cap="Exemplary simulation: Gaussian kernel density estimates of the marginal posterior distributions for setup 3", fig.width=6, fig.height=6.5, warning=FALSE}
traces <- sim_4$pmmh_false_informative$traces[10000:20000, ]

# Plot theme:
p_theme <- theme(
  legend.justification = c(0, 1), legend.position = c(0, 1), 
  legend.title = element_blank(), 
  legend.background = element_rect(fill = "transparent"),
  plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
  plot.title = element_text(hjust = 0.5, size = 8),
  axis.title.x = element_text(size = 8),
  axis.title.y = element_text(size = 8, vjust = 0.5),
  legend.key = element_rect(fill = NA)
)

# Beta 0
df <- data.frame(time = 1:nrow(traces), beta_0 = traces[, 3])
p1 <- ggplot(data = df, aes(x = beta_0)) +
  geom_density(aes(x = beta_0, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = true_params_simulation["beta_0"], colour = "True",
        linetype = "True"), show.legend = FALSE
  ) +
  geom_vline(
    aes(xintercept = mean(traces[, 3]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = dnorm, aes(colour = "Prior", linetype = "Prior"),
    args = list(mean = 0.8, sd = 0.1)
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", "True" = "red",
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, 
               "Prior" = 1, "True" = 3), name = ""
  ) + 
  labs(title = latex2exp::TeX("Marginal Posterior: \\textbf{$\\beta_0$}"),
       x = latex2exp::TeX("$\\beta_0$"))

# Beta 1
df <- data.frame(time = 1:nrow(traces), beta_1 = traces[, 4])
p2 <- ggplot(data = df, aes(x = beta_1)) +
  geom_density(aes(x = beta_1, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = true_params_simulation["beta_1"], colour = "True",
        linetype = "True"), show.legend = FALSE
  ) +
  geom_vline(
    aes(xintercept = mean(traces[, 4]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = dnorm, aes(colour = "Prior", linetype = "Prior"),
    args = list(mean = 0.0, sd = 0.1)
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", "True" = "red",
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, 
               "Prior" = 1, "True" = 3), name = ""
  ) + 
  labs(title = latex2exp::TeX("Marginal Posterior: \\textbf{$\\beta_1$}"),
       x = latex2exp::TeX("$\\beta_1$"))

# Beta 2
df <- data.frame(time = 1:nrow(traces), beta_2 = traces[, 5])
p3 <- ggplot(data = df, aes(x = beta_2)) +
  geom_density(aes(x = beta_2, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = true_params_simulation["beta_2"], colour = "True",
        linetype = "True"), show.legend = FALSE
  ) +
  geom_vline(
    aes(xintercept = mean(traces[, 5]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = dnorm, aes(colour = "Prior", linetype = "Prior"),
    args = list(mean = 0.0, sd = 0.1)
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", "True" = "red",
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, 
               "Prior" = 1, "True" = 3), name = ""
  ) + 
  labs(title = latex2exp::TeX("Marginal Posterior: \\textbf{$\\beta_2$}"),
       x = latex2exp::TeX("$\\beta_2$"))

# Sigma x
df <- data.frame(time = 1:nrow(traces), sigma_x = exp(traces[, 6]))
p4 <- ggplot(data = df, aes(x = sigma_x)) +
  geom_density(aes(x = sigma_x, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = exp(true_params_simulation["sigma_x"]), colour = "True",
        linetype = "True"), show.legend = FALSE
  ) +
  geom_vline(
    aes(xintercept = mean(exp(traces[, 6])), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = dgamma, aes(colour = "Prior", linetype = "Prior"),
    args = list(shape = 10, scale = 0.01)
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", "True" = "red",
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, 
               "Prior" = 1, "True" = 3), name = ""
  ) + 
  labs(title = latex2exp::TeX("Marginal Posterior: \\textbf{$\\sigma_u$}"),
       x = latex2exp::TeX("$\\sigma_u$"))

# Sigma y
df <- data.frame(time = 1:nrow(traces), sigma_y = exp(traces[, 7]))
p5 <- ggplot(data = df, aes(x = sigma_y)) +
  geom_density(aes(x = sigma_y, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = exp(true_params_simulation["sigma_y"]), colour = "True",
        linetype = "True"), show.legend = FALSE
  ) +
  geom_vline(
    aes(xintercept = mean(exp(traces[, 7])), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = dgamma, aes(colour = "Prior", linetype = "Prior"),
    args = list(shape = 10, scale = 0.005)
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", "True" = "red",
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, 
               "Prior" = 1, "True" = 3), name = ""
  ) + 
  labs(title = latex2exp::TeX("Marginal Posterior: \\textbf{$\\sigma_\\epsilon$}"),
       x = latex2exp::TeX("$\\sigma_\\epsilon$"))

# Phi
df <- data.frame(time = 1:nrow(traces), phi = tanh(traces[, 8]))
p6 <- ggplot(data = df, aes(x = phi)) +
  geom_density(aes(x = phi, colour = "Posterior", linetype = "Posterior"),
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = tanh(true_params_simulation["phi"]), colour = "True",
        linetype = "True"), show.legend = FALSE
  ) +
  geom_vline(
    aes(xintercept = mean(tanh(traces[, 8])), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = truncnorm::dtruncnorm, aes(colour = "Prior", linetype = "Prior"),
    args = list(a = -1, b = 1, mean = 0.3, sd = 0.1)
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", "True" = "red",
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, 
               "Prior" = 1, "True" = 3), name = ""
  ) + 
  labs(title = latex2exp::TeX("Marginal Posterior: \\textbf{$\\phi$}"),
       x = latex2exp::TeX("$\\phi$"))

# X 0
df <- data.frame(time = 1:nrow(traces), x_0 = traces[, 9])
p7 <- ggplot(data = df, aes(x = x_0)) +
  geom_density(aes(x = x_0, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = true_params_simulation["x_0"], colour = "True",
        linetype = "True"), show.legend = FALSE
  ) +
  geom_vline(
    aes(xintercept = mean(traces[, 9]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = dnorm, aes(colour = "Prior", linetype = "Prior"),
    args = list(mean = 3.0, sd = 0.1)
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", "True" = "red",
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, 
               "Prior" = 1, "True" = 3), name = ""
  ) + 
  labs(title = latex2exp::TeX("Marginal Posterior: \\textbf{$\\x_0$}"),
       x = latex2exp::TeX("$\\x_0$"))

get_legend <- function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}
legend <- get_legend(p2)

grid.arrange(
  p1 + theme(legend.position = "none"), 
  p2 + theme(legend.position = "none"), 
  p3 + theme(legend.position = "none"), 
  p4 + theme(legend.position = "none"), 
  p5 + theme(legend.position = "none"), 
  p6 + theme(legend.position = "none"), 
  p7 + theme(legend.position = "none"), 
  ncol = 2, legend = legend
)
rm(list = ls())
```

```{r, extension-static-model-setup, eval=TRUE, echo=FALSE}
# get simulation result
source(here::here("R", "model_comparison", "model_comparison_static_model.R"))
library(ggplot2)
library(gridExtra)
```

```{r, extension-static-model-trace, eval=TRUE, echo=FALSE, fig.align='center', fig.cap="Covariate extension: Trace plot setup 3", fig.width=7, fig.height=8.5, warning=FALSE}
traces <- list(
  result[[1]]@traces, 
  result[[2]]@traces, 
  result[[3]]@traces
)

# Plot theme:
p_theme <- theme(
  legend.position = "bottom", 
  legend.title = element_blank(), 
  legend.background = element_rect(fill = "transparent"),
  plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
  plot.title = element_text(hjust = 0.5, size = 8),
  axis.title.x = element_text(size = 8),
  axis.title.y = element_text(size = 8, angle = 0, vjust = 0.5)
)

# beta_0
df <- data.frame(
  time = 1:nrow(traces[[1]]),
  value1 = traces[[1]][, 3], 
  value2 = traces[[2]][, 3], 
  value3 = traces[[3]][, 3]
)
df <- reshape2::melt(df, id.vars = "time")
p1 <- ggplot(data = df, aes(x = time, y = value)) +
  geom_line(aes(colour = variable)) + theme_light() +
  scale_colour_manual(
    values = c("red", "green", "blue"),
    labels = c("Chain 1", "Chain 2", "Chain 3")
  ) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\beta_0$}"),
       x = "m", y = latex2exp::TeX("$\\beta_0$")) +
  p_theme

# beta_1
df <- data.frame(
  time = 1:nrow(traces[[1]]),
  value1 = traces[[1]][, 4], 
  value2 = traces[[2]][, 4], 
  value3 = traces[[3]][, 4]
)
df <- reshape2::melt(df, id.vars = "time")
p2 <- ggplot(data = df, aes(x = time, y = value)) +
  geom_line(aes(colour = variable)) + theme_light() +
  scale_colour_manual(
    values = c("red", "green", "blue"),
    labels = c("Chain 1", "Chain 2", "Chain 3")
  ) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\beta_1$}"),
       x = "m", y = latex2exp::TeX("$\\beta_1$")) +
  p_theme

# beta_2
df <- data.frame(
  time = 1:nrow(traces[[1]]),
  value1 = traces[[1]][, 5], 
  value2 = traces[[2]][, 5], 
  value3 = traces[[3]][, 5]
)
df <- reshape2::melt(df, id.vars = "time")
p3 <- ggplot(data = df, aes(x = time, y = value)) +
  geom_line(aes(colour = variable)) + theme_light() +
  scale_colour_manual(
    values = c("red", "green", "blue"),
    labels = c("Chain 1", "Chain 2", "Chain 3")
  ) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\beta_2$}"),
       x = "m", y = latex2exp::TeX("$\\beta_2$")) +
  p_theme

# beta_3
df <- data.frame(
  time = 1:nrow(traces[[1]]),
  value1 = traces[[1]][, 6], 
  value2 = traces[[2]][, 6], 
  value3 = traces[[3]][, 6]
)
df <- reshape2::melt(df, id.vars = "time")
p4 <- ggplot(data = df, aes(x = time, y = value)) +
  geom_line(aes(colour = variable)) + theme_light() +
  scale_colour_manual(
    values = c("red", "green", "blue"),
    labels = c("Chain 1", "Chain 2", "Chain 3")
  ) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\beta_3$}"),
       x = "m", y = latex2exp::TeX("$\\beta_3$")) +
  p_theme

# beta_4
df <- data.frame(
  time = 1:nrow(traces[[1]]),
  value1 = traces[[1]][, 7], 
  value2 = traces[[2]][, 7], 
  value3 = traces[[3]][, 7]
)
df <- reshape2::melt(df, id.vars = "time")
p5 <- ggplot(data = df, aes(x = time, y = value)) +
  geom_line(aes(colour = variable)) + theme_light() +
  scale_colour_manual(
    values = c("red", "green", "blue"),
    labels = c("Chain 1", "Chain 2", "Chain 3")
  ) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\beta_4$}"),
       x = "m", y = latex2exp::TeX("$\\beta_4$")) +
  p_theme

# beta_5
df <- data.frame(
  time = 1:nrow(traces[[1]]),
  value1 = traces[[1]][, 8], 
  value2 = traces[[2]][, 8], 
  value3 = traces[[3]][, 8]
)
df <- reshape2::melt(df, id.vars = "time")
p6 <- ggplot(data = df, aes(x = time, y = value)) +
  geom_line(aes(colour = variable)) + theme_light() +
  scale_colour_manual(
    values = c("red", "green", "blue"),
    labels = c("Chain 1", "Chain 2", "Chain 3")
  ) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\beta_5$}"),
       x = "m", y = latex2exp::TeX("$\\beta_5$")) +
  p_theme

# sigma_u
df <- data.frame(
  time = 1:nrow(traces[[1]]),
  value1 = traces[[1]][, 9], 
  value2 = traces[[2]][, 9], 
  value3 = traces[[3]][, 9]
)
df <- reshape2::melt(df, id.vars = "time")
df[, "value"] <- exp(df[, "value"])
p7 <- ggplot(data = df, aes(x = time, y = value)) +
  geom_line(aes(colour = variable)) + theme_light() +
  scale_colour_manual(
    values = c("red", "green", "blue"),
    labels = c("Chain 1", "Chain 2", "Chain 3")
  ) +
  labs(title = latex2exp::TeX("Trace Plot: $\\sigma_u$"),
       x = "m", y = latex2exp::TeX("$\\sigma_u$")) +
  p_theme

# sigma_u
df <- data.frame(
  time = 1:nrow(traces[[1]]),
  value1 = traces[[1]][, 10], 
  value2 = traces[[2]][, 10], 
  value3 = traces[[3]][, 10]
)
df <- reshape2::melt(df, id.vars = "time")
df[, "value"] <- exp(df[, "value"])
p8 <- ggplot(data = df, aes(x = time, y = value)) +
  geom_line(aes(colour = variable)) + theme_light() +
  scale_colour_manual(
    values = c("red", "green", "blue"),
    labels = c("Chain 1", "Chain 2", "Chain 3")
  ) +
  labs(title = latex2exp::TeX("Trace Plot: $\\sigma_\\epsilon$"),
       x = "m", y = latex2exp::TeX("$\\sigma_\\epsilon$")) +
  p_theme

legend <- gtable::gtable_filter(ggplot_gtable(ggplot_build(p2)), "guide-box")

grid.arrange(
  arrangeGrob(
    p1 + theme(legend.position = "none"), 
    p2 + theme(legend.position = "none"), 
    p3 + theme(legend.position = "none"), 
    p4 + theme(legend.position = "none"), 
    p5 + theme(legend.position = "none"), 
    p6 + theme(legend.position = "none"), 
    p7 + theme(legend.position = "none"), 
    p8 + theme(legend.position = "none"),
    ncol = 2
  ),
  arrangeGrob(grid::nullGrob(), legend, grid::nullGrob(), nrow = 1),
  heights = c(4.5, 0.1), ncol = 1
)
```

```{r, extension-static-model-marginal-density, eval=TRUE, echo=FALSE, fig.align='center', fig.cap="Random walk state model: Gaussian kernel density estimates of the marginal posterior distributions", fig.width=7, fig.height=8.5, warning=FALSE}
traces <- rbind(
  result[[1]]@traces[30000:50000, ], 
  result[[2]]@traces[30000:50000, ], 
  result[[3]]@traces[30000:50000, ]
)

# Plot theme:
p_theme <- theme(
  legend.position = "bottom",
  legend.title = element_blank(), 
  legend.background = element_rect(fill = "transparent"),
  plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
  plot.title = element_text(hjust = 0.5, size = 8),
  axis.title.x = element_text(size = 8),
  axis.title.y = element_text(size = 8, vjust = 0.5)
)

# beta_0
df <- data.frame(time = 1:nrow(traces), beta_0 = traces[, 3])
p1 <- ggplot(data = df, aes(x = beta_0)) +
  geom_density(aes(x = beta_0, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = mean(traces[, 3]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", 
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, "Prior" = 1), name = ""
  ) + 
  labs(title = latex2exp::TeX(
    "Marginal Posterior: \\textbf{$\\beta_0$}"
  ), x = latex2exp::TeX("$\\beta_0$"))

# beta_1
df <- data.frame(time = 1:nrow(traces), beta_1 = traces[, 4])
p2 <- ggplot(data = df, aes(x = beta_1)) +
  geom_density(aes(x = beta_1, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = mean(traces[, 4]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", 
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, "Prior" = 1), name = ""
  ) + 
  labs(title = latex2exp::TeX(
    "Marginal Posterior: \\textbf{$\\beta_1$}"
  ), x = latex2exp::TeX("$\\beta_1$"))

# beta_2
df <- data.frame(time = 1:nrow(traces), beta_2 = traces[, 5])
p3 <- ggplot(data = df, aes(x = beta_2)) +
  geom_density(aes(x = beta_2, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = mean(traces[, 5]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", 
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, "Prior" = 1), name = ""
  ) + 
  labs(title = latex2exp::TeX(
    "Marginal Posterior: \\textbf{$\\beta_2$}"
  ), x = latex2exp::TeX("$\\beta_2$"))

# beta_3
df <- data.frame(time = 1:nrow(traces), beta_3 = traces[, 6])
p4 <- ggplot(data = df, aes(x = beta_3)) +
  geom_density(aes(x = beta_3, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = mean(traces[, 6]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", 
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, "Prior" = 1), name = ""
  ) + 
  labs(title = latex2exp::TeX(
    "Marginal Posterior: \\textbf{$\\beta_3$}"
  ), x = latex2exp::TeX("$\\beta_3$"))

# beta_4
df <- data.frame(time = 1:nrow(traces), beta_4 = traces[, 7])
p5 <- ggplot(data = df, aes(x = beta_4)) +
  geom_density(aes(x = beta_4, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = mean(traces[, 7]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", 
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, "Prior" = 1), name = ""
  ) + 
  labs(title = latex2exp::TeX(
    "Marginal Posterior: \\textbf{$\\beta_4$}"
  ), x = latex2exp::TeX("$\\beta_4$"))

# beta_5
df <- data.frame(time = 1:nrow(traces), beta_5 = traces[, 8])
p6 <- ggplot(data = df, aes(x = beta_5)) +
  geom_density(aes(x = beta_5, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = mean(traces[, 8]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", 
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, "Prior" = 1), name = ""
  ) + 
  labs(title = latex2exp::TeX(
    "Marginal Posterior: \\textbf{$\\beta_5$}"
  ), x = latex2exp::TeX("$\\beta_5$"))

# sigma_u
df <- data.frame(time = 1:nrow(traces), sigma_u = exp(traces[, 9]))
p7 <- ggplot(data = df, aes(x = sigma_u)) +
  geom_density(aes(x = sigma_u, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = mean(exp(traces[, 9])), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", 
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, "Prior" = 1), name = ""
  ) + 
  labs(title = latex2exp::TeX("Marginal Posterior: \\textbf{$\\sigma_u$}"),
       x = latex2exp::TeX("$\\sigma_u$"))

# sigma_epsilon
df <- data.frame(time = 1:nrow(traces), sigma_e = exp(traces[, 10]))
p8 <- ggplot(data = df, aes(x = sigma_e)) +
  geom_density(aes(x = sigma_e, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = mean(exp(traces[, 10])), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", 
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, "Prior" = 1), name = ""
  ) + 
  labs(title = latex2exp::TeX(
    "Marginal Posterior: \\textbf{$\\sigma_\\epsilon$}"
  ), x = latex2exp::TeX("$\\sigma_\\epsilon$"))

get_legend <- function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}
legend <- get_legend(p2)

grid.arrange(
  arrangeGrob(
    p1 + theme(legend.position = "none"), 
    p2 + theme(legend.position = "none"), 
    p3 + theme(legend.position = "none"), 
    p4 + theme(legend.position = "none"), 
    p5 + theme(legend.position = "none"), 
    p6 + theme(legend.position = "none"), 
    p7 + theme(legend.position = "none"), 
    p8 + theme(legend.position = "none"),
    ncol = 2
  ),
  arrangeGrob(grid::nullGrob(), legend, grid::nullGrob(), nrow = 1),
  heights = c(4.5, 0.1), ncol = 1
)
```

```{r, extension-ar-error-model-setup, eval=TRUE, echo=FALSE}
# get simulation result
source(here::here("R", "model_comparison", "model_comparison_ar_error_model.R"))
library(ggplot2)
library(gridExtra)
```

```{r, extension-ar-error-model-trace, eval=TRUE, echo=FALSE, fig.align='center', fig.cap="Covariate extension: Trace plot setup 3", fig.width=6, fig.height=9.5, warning=FALSE}
traces <- list(
  result[[1]]@traces, 
  result[[2]]@traces, 
  result[[3]]@traces
)

# Plot theme:
p_theme <- theme(
  legend.justification = c(0, 1), legend.position = c(0, 1), 
  legend.title = element_blank(), 
  legend.background = element_rect(fill = "transparent"),
  plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
  plot.title = element_text(hjust = 0.5, size = 8),
  axis.title.x = element_text(size = 8),
  axis.title.y = element_text(size = 8, angle = 0, vjust = 0.5)
)

# beta_0
df <- data.frame(
  time = 1:nrow(traces[[1]]),
  value1 = traces[[1]][, 7], 
  value2 = traces[[2]][, 7], 
  value3 = traces[[3]][, 7]
)
df <- reshape2::melt(df, id.vars = "time")
p1 <- ggplot(data = df, aes(x = time, y = value)) +
  geom_line(aes(colour = variable)) + theme_light() +
  scale_colour_manual(
    values = c("red", "green", "blue"),
    labels = c("Chain 1", "Chain 2", "Chain 3")
  ) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\beta_0$}"),
       x = "m", y = latex2exp::TeX("$\\beta_0$")) +
  p_theme

# beta_1
df <- data.frame(
  time = 1:nrow(traces[[1]]),
  value1 = traces[[1]][, 8], 
  value2 = traces[[2]][, 8], 
  value3 = traces[[3]][, 8]
)
df <- reshape2::melt(df, id.vars = "time")
p2 <- ggplot(data = df, aes(x = time, y = value)) +
  geom_line(aes(colour = variable)) + theme_light() +
  scale_colour_manual(
    values = c("red", "green", "blue"),
    labels = c("Chain 1", "Chain 2", "Chain 3")
  ) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\beta_1$}"),
       x = "m", y = latex2exp::TeX("$\\beta_1$")) +
  p_theme

# beta_2
df <- data.frame(
  time = 1:nrow(traces[[1]]),
  value1 = traces[[1]][, 9], 
  value2 = traces[[2]][, 9], 
  value3 = traces[[3]][, 9]
)
df <- reshape2::melt(df, id.vars = "time")
p3 <- ggplot(data = df, aes(x = time, y = value)) +
  geom_line(aes(colour = variable)) + theme_light() +
  scale_colour_manual(
    values = c("red", "green", "blue"),
    labels = c("Chain 1", "Chain 2", "Chain 3")
  ) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\beta_2$}"),
       x = "m", y = latex2exp::TeX("$\\beta_2$")) +
  p_theme

# beta_3
df <- data.frame(
  time = 1:nrow(traces[[1]]),
  value1 = traces[[1]][, 10], 
  value2 = traces[[2]][, 10], 
  value3 = traces[[3]][, 10]
)
df <- reshape2::melt(df, id.vars = "time")
p4 <- ggplot(data = df, aes(x = time, y = value)) +
  geom_line(aes(colour = variable)) + theme_light() +
  scale_colour_manual(
    values = c("red", "green", "blue"),
    labels = c("Chain 1", "Chain 2", "Chain 3")
  ) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\beta_3$}"),
       x = "m", y = latex2exp::TeX("$\\beta_3$")) +
  p_theme

# beta_4
df <- data.frame(
  time = 1:nrow(traces[[1]]),
  value1 = traces[[1]][, 11], 
  value2 = traces[[2]][, 11], 
  value3 = traces[[3]][, 11]
)
df <- reshape2::melt(df, id.vars = "time")
p5 <- ggplot(data = df, aes(x = time, y = value)) +
  geom_line(aes(colour = variable)) + theme_light() +
  scale_colour_manual(
    values = c("red", "green", "blue"),
    labels = c("Chain 1", "Chain 2", "Chain 3")
  ) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\beta_4$}"),
       x = "m", y = latex2exp::TeX("$\\beta_4$")) +
  p_theme

# beta_5
df <- data.frame(
  time = 1:nrow(traces[[1]]),
  value1 = traces[[1]][, 12], 
  value2 = traces[[2]][, 12], 
  value3 = traces[[3]][, 12]
)
df <- reshape2::melt(df, id.vars = "time")
p6 <- ggplot(data = df, aes(x = time, y = value)) +
  geom_line(aes(colour = variable)) + theme_light() +
  scale_colour_manual(
    values = c("red", "green", "blue"),
    labels = c("Chain 1", "Chain 2", "Chain 3")
  ) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\beta_5$}"),
       x = "m", y = latex2exp::TeX("$\\beta_5$")) +
  p_theme

# sigma_u
df <- data.frame(
  time = 1:nrow(traces[[1]]),
  value1 = traces[[1]][, 5], 
  value2 = traces[[2]][, 5], 
  value3 = traces[[3]][, 5]
)
df <- reshape2::melt(df, id.vars = "time")
df[, "value"] <- exp(df[, "value"])
p7 <- ggplot(data = df, aes(x = time, y = value)) +
  geom_line(aes(colour = variable)) + theme_light() +
  scale_colour_manual(
    values = c("red", "green", "blue"),
    labels = c("Chain 1", "Chain 2", "Chain 3")
  ) +
  labs(title = latex2exp::TeX("Trace Plot: $\\sigma_u$"),
       x = "m", y = latex2exp::TeX("$\\sigma_u$")) +
  p_theme

# sigma_e
df <- data.frame(
  time = 1:nrow(traces[[1]]),
  value1 = traces[[1]][, 4], 
  value2 = traces[[2]][, 4], 
  value3 = traces[[3]][, 4]
)
df <- reshape2::melt(df, id.vars = "time")
df[, "value"] <- exp(df[, "value"])
p8 <- ggplot(data = df, aes(x = time, y = value)) +
  geom_line(aes(colour = variable)) + theme_light() +
  scale_colour_manual(
    values = c("red", "green", "blue"),
    labels = c("Chain 1", "Chain 2", "Chain 3")
  ) +
  labs(title = latex2exp::TeX("Trace Plot: $\\sigma_\\epsilon$"),
       x = "m", y = latex2exp::TeX("$\\sigma_\\epsilon$")) +
  p_theme

# phi
df <- data.frame(
  time = 1:nrow(traces[[1]]),
  value1 = traces[[1]][, 6], 
  value2 = traces[[2]][, 6], 
  value3 = traces[[3]][, 6]
)
df <- reshape2::melt(df, id.vars = "time")
df[, "value"] <- tanh(df[, "value"])
p9 <- ggplot(data = df, aes(x = time, y = value)) +
  geom_line(aes(colour = variable)) + theme_light() +
  scale_colour_manual(
    values = c("red", "green", "blue"),
    labels = c("Chain 1", "Chain 2", "Chain 3")
  ) +
  labs(title = latex2exp::TeX("Trace Plot: $\\phi$"),
       x = "m", y = latex2exp::TeX("$\\phi$")) +
  p_theme

legend <- gtable::gtable_filter(ggplot_gtable(ggplot_build(p2)), "guide-box")

grid.arrange(
  p1 + theme(legend.position = "none"), 
  p2 + theme(legend.position = "none"), 
  p3 + theme(legend.position = "none"), 
  p4 + theme(legend.position = "none"), 
  p5 + theme(legend.position = "none"), 
  p6 + theme(legend.position = "none"), 
  p7 + theme(legend.position = "none"), 
  p8 + theme(legend.position = "none"),
  p9 + theme(legend.position = "none"),
  legend = legend,
  heights = c(rep(1.1, 5)), ncol = 2, nrow = 5
)
```

```{r, extension-ar-error-model-marginal-density, eval=TRUE, echo=FALSE, fig.align='center', fig.cap="Random walk state model: Gaussian kernel density estimates of the marginal posterior distributions", fig.width=6, fig.height=8.5, warning=FALSE}
traces <- rbind(
  result[[1]]@traces[5000:20000, ], 
  result[[2]]@traces[5000:20000, ], 
  result[[3]]@traces[5000:20000, ]
)

# Plot theme:
p_theme <- theme(
  legend.justification = c(0, 1), legend.position = c(0, 1), 
  legend.title = element_blank(), 
  legend.background = element_rect(fill = "transparent"),
  plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
  plot.title = element_text(hjust = 0.5, size = 8),
  axis.title.x = element_text(size = 8),
  axis.title.y = element_text(size = 8, vjust = 0.5)
)

# beta_0
df <- data.frame(time = 1:nrow(traces), beta_0 = traces[, 6])
p1 <- ggplot(data = df, aes(x = beta_0)) +
  geom_density(aes(x = beta_0, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = mean(traces[, 6]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", 
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, "Prior" = 1), name = ""
  ) + 
  labs(title = latex2exp::TeX(
    "Marginal Posterior: \\textbf{$\\beta_0$}"
  ), x = latex2exp::TeX("$\\beta_0$"))

# beta_1
df <- data.frame(time = 1:nrow(traces), beta_1 = traces[, 7])
p2 <- ggplot(data = df, aes(x = beta_1)) +
  geom_density(aes(x = beta_1, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = mean(traces[, 7]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", 
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, "Prior" = 1), name = ""
  ) + 
  labs(title = latex2exp::TeX(
    "Marginal Posterior: \\textbf{$\\beta_1$}"
  ), x = latex2exp::TeX("$\\beta_1$"))

# beta_2
df <- data.frame(time = 1:nrow(traces), beta_2 = traces[, 8])
p3 <- ggplot(data = df, aes(x = beta_2)) +
  geom_density(aes(x = beta_2, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = mean(traces[, 8]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", 
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, "Prior" = 1), name = ""
  ) + 
  labs(title = latex2exp::TeX(
    "Marginal Posterior: \\textbf{$\\beta_2$}"
  ), x = latex2exp::TeX("$\\beta_2$"))

# beta_3
df <- data.frame(time = 1:nrow(traces), beta_3 = traces[, 9])
p4 <- ggplot(data = df, aes(x = beta_3)) +
  geom_density(aes(x = beta_3, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = mean(traces[, 9]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", 
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, "Prior" = 1), name = ""
  ) + 
  labs(title = latex2exp::TeX(
    "Marginal Posterior: \\textbf{$\\beta_3$}"
  ), x = latex2exp::TeX("$\\beta_3$"))

# beta_4
df <- data.frame(time = 1:nrow(traces), beta_4 = traces[, 10])
p5 <- ggplot(data = df, aes(x = beta_4)) +
  geom_density(aes(x = beta_4, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = mean(traces[, 10]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", 
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, "Prior" = 1), name = ""
  ) + 
  labs(title = latex2exp::TeX(
    "Marginal Posterior: \\textbf{$\\beta_4$}"
  ), x = latex2exp::TeX("$\\beta_4$"))

# beta_5
df <- data.frame(time = 1:nrow(traces), beta_5 = traces[, 11])
p6 <- ggplot(data = df, aes(x = beta_5)) +
  geom_density(aes(x = beta_5, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = mean(traces[, 11]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", 
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, "Prior" = 1), name = ""
  ) + 
  labs(title = latex2exp::TeX(
    "Marginal Posterior: \\textbf{$\\beta_5$}"
  ), x = latex2exp::TeX("$\\beta_5$"))

# sigma_u
df <- data.frame(time = 1:nrow(traces), sigma_u = exp(traces[, 4]))
p7 <- ggplot(data = df, aes(x = sigma_u)) +
  geom_density(aes(x = sigma_u, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = mean(exp(traces[, 4])), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", 
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, "Prior" = 1), name = ""
  ) + 
  labs(title = latex2exp::TeX("Marginal Posterior: \\textbf{$\\sigma_u$}"),
       x = latex2exp::TeX("$\\sigma_u$"))

# sigma_epsilon
df <- data.frame(time = 1:nrow(traces), sigma_e = exp(traces[, 3]))
p8 <- ggplot(data = df, aes(x = sigma_e)) +
  geom_density(aes(x = sigma_e, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = mean(exp(traces[, 3])), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", 
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, "Prior" = 1), name = ""
  ) + 
  labs(title = latex2exp::TeX(
    "Marginal Posterior: \\textbf{$\\sigma_\\epsilon$}"
  ), x = latex2exp::TeX("$\\sigma_\\epsilon$"))

# phi
df <- data.frame(time = 1:nrow(traces), phi = tanh(traces[, 5]))
p9 <- ggplot(data = df, aes(x = phi)) +
  geom_density(aes(x = phi, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = mean(tanh(traces[, 5])), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", 
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, "Prior" = 1), name = ""
  ) + 
  labs(title = latex2exp::TeX(
    "Marginal Posterior: \\textbf{$\\phi$}"
  ), x = latex2exp::TeX("$\\phi$"))

get_legend <- function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}
legend <- get_legend(p2)

grid.arrange(
  p1 + theme(legend.position = "none"), 
  p2 + theme(legend.position = "none"), 
  p3 + theme(legend.position = "none"), 
  p4 + theme(legend.position = "none"), 
  p5 + theme(legend.position = "none"), 
  p6 + theme(legend.position = "none"), 
  p7 + theme(legend.position = "none"), 
  p8 + theme(legend.position = "none"),
  p9 + theme(legend.position = "none"),
  legend = legend,
  heights = rep(1.1, 5), ncol = 2
)
```

```{r, extension-ardl-model-setup, eval=FALSE, echo=FALSE}
# get simulation result
source(here::here("R", "model_comparison", "model_comparison_ardl_model.R"))
library(ggplot2)
library(gridExtra)
```

```{r, extension-ardl-model-trace, eval=FALSE, echo=FALSE, fig.align='center', fig.cap="Covariate extension: Trace plot setup 3", fig.width=6, fig.height=8.5, warning=FALSE}
traces <- list(
  result[[1]]@traces, 
  result[[2]]@traces, 
  result[[3]]@traces
)

# Plot theme:
p_theme <- theme(
  legend.justification = c(0, 1), legend.position = c(0, 1), 
  legend.title = element_blank(), 
  legend.background = element_rect(fill = "transparent"),
  plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
  plot.title = element_text(hjust = 0.5, size = 8),
  axis.title.x = element_text(size = 8),
  axis.title.y = element_text(size = 8, angle = 0, vjust = 0.5)
)

# e_lpd_0
df <- data.frame(
  time = 1:nrow(traces[[1]]),
  value1 = traces[[1]][, 3], 
  value2 = traces[[2]][, 3], 
  value3 = traces[[3]][, 3]
)
df <- reshape2::melt(df, id.vars = "time")
p1 <- ggplot(data = df, aes(x = time, y = value)) +
  geom_line(aes(colour = variable)) + theme_light() +
  scale_colour_manual(
    values = c("red", "green", "blue"),
    labels = c("Chain 1", "Chain 2", "Chain 3")
  ) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\tilde{\\eta}_0$}"),
       x = "m", y = latex2exp::TeX("$\\tilde{\\eta}_0$")) +
  p_theme

# beta_0
df <- data.frame(
  time = 1:nrow(traces[[1]]),
  value1 = traces[[1]][, 7], 
  value2 = traces[[2]][, 7], 
  value3 = traces[[3]][, 7]
)
df <- reshape2::melt(df, id.vars = "time")
p2 <- ggplot(data = df, aes(x = time, y = value)) +
  geom_line(aes(colour = variable)) + theme_light() +
  scale_colour_manual(
    values = c("red", "green", "blue"),
    labels = c("Chain 1", "Chain 2", "Chain 3")
  ) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\beta_0$}"),
       x = "m", y = latex2exp::TeX("$\\beta_0$")) +
  p_theme

# beta_1
df <- data.frame(
  time = 1:nrow(traces[[1]]),
  value1 = traces[[1]][, 8], 
  value2 = traces[[2]][, 8], 
  value3 = traces[[3]][, 8]
)
df <- reshape2::melt(df, id.vars = "time")
p3 <- ggplot(data = df, aes(x = time, y = value)) +
  geom_line(aes(colour = variable)) + theme_light() +
  scale_colour_manual(
    values = c("red", "green", "blue"),
    labels = c("Chain 1", "Chain 2", "Chain 3")
  ) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\beta_1$}"),
       x = "m", y = latex2exp::TeX("$\\beta_1$")) +
  p_theme

# beta_2
df <- data.frame(
  time = 1:nrow(traces[[1]]),
  value1 = traces[[1]][, 9], 
  value2 = traces[[2]][, 9], 
  value3 = traces[[3]][, 9]
)
df <- reshape2::melt(df, id.vars = "time")
p4 <- ggplot(data = df, aes(x = time, y = value)) +
  geom_line(aes(colour = variable)) + theme_light() +
  scale_colour_manual(
    values = c("red", "green", "blue"),
    labels = c("Chain 1", "Chain 2", "Chain 3")
  ) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\beta_2$}"),
       x = "m", y = latex2exp::TeX("$\\beta_2$")) +
  p_theme

# beta_3
df <- data.frame(
  time = 1:nrow(traces[[1]]),
  value1 = traces[[1]][, 10], 
  value2 = traces[[2]][, 10], 
  value3 = traces[[3]][, 10]
)
df <- reshape2::melt(df, id.vars = "time")
p5 <- ggplot(data = df, aes(x = time, y = value)) +
  geom_line(aes(colour = variable)) + theme_light() +
  scale_colour_manual(
    values = c("red", "green", "blue"),
    labels = c("Chain 1", "Chain 2", "Chain 3")
  ) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\beta_3$}"),
       x = "m", y = latex2exp::TeX("$\\beta_3$")) +
  p_theme

# beta_4
df <- data.frame(
  time = 1:nrow(traces[[1]]),
  value1 = traces[[1]][, 11], 
  value2 = traces[[2]][, 11], 
  value3 = traces[[3]][, 11]
)
df <- reshape2::melt(df, id.vars = "time")
p6 <- ggplot(data = df, aes(x = time, y = value)) +
  geom_line(aes(colour = variable)) + theme_light() +
  scale_colour_manual(
    values = c("red", "green", "blue"),
    labels = c("Chain 1", "Chain 2", "Chain 3")
  ) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\beta_4$}"),
       x = "m", y = latex2exp::TeX("$\\beta_4$")) +
  p_theme

# beta_5
df <- data.frame(
  time = 1:nrow(traces[[1]]),
  value1 = traces[[1]][, 12], 
  value2 = traces[[2]][, 12], 
  value3 = traces[[3]][, 12]
)
df <- reshape2::melt(df, id.vars = "time")
p7 <- ggplot(data = df, aes(x = time, y = value)) +
  geom_line(aes(colour = variable)) + theme_light() +
  scale_colour_manual(
    values = c("red", "green", "blue"),
    labels = c("Chain 1", "Chain 2", "Chain 3")
  ) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\beta_5$}"),
       x = "m", y = latex2exp::TeX("$\\beta_5$")) +
  p_theme

# phi
df <- data.frame(
  time = 1:nrow(traces[[1]]),
  value1 = traces[[1]][, 6], 
  value2 = traces[[2]][, 6], 
  value3 = traces[[3]][, 6]
)
df <- reshape2::melt(df, id.vars = "time")
df[, "value"] <- tanh(df[, "value"])
p8 <- ggplot(data = df, aes(x = time, y = value)) +
  geom_line(aes(colour = variable)) + theme_light() +
  scale_colour_manual(
    values = c("red", "green", "blue"),
    labels = c("Chain 1", "Chain 2", "Chain 3")
  ) +
  labs(title = latex2exp::TeX("Trace Plot: \\textbf{$\\phi$}"),
       x = "m", y = latex2exp::TeX("$\\phi$")) +
  p_theme

# sigma_u
df <- data.frame(
  time = 1:nrow(traces[[1]]),
  value1 = traces[[1]][, 5], 
  value2 = traces[[2]][, 5], 
  value3 = traces[[3]][, 5]
)
df <- reshape2::melt(df, id.vars = "time")
df[, "value"] <- exp(df[, "value"])
p9 <- ggplot(data = df, aes(x = time, y = value)) +
  geom_line(aes(colour = variable)) + theme_light() +
  scale_colour_manual(
    values = c("red", "green", "blue"),
    labels = c("Chain 1", "Chain 2", "Chain 3")
  ) +
  labs(title = latex2exp::TeX("Trace Plot: $\\sigma_u$"),
       x = "m", y = latex2exp::TeX("$\\sigma_u$")) +
  p_theme

# sigma_epsilon
df <- data.frame(
  time = 1:nrow(traces[[1]]),
  value1 = traces[[1]][, 4], 
  value2 = traces[[2]][, 4], 
  value3 = traces[[3]][, 4]
)
df <- reshape2::melt(df, id.vars = "time")
df[, "value"] <- exp(df[, "value"])
p10 <- ggplot(data = df, aes(x = time, y = value)) +
  geom_line(aes(colour = variable)) + theme_light() +
  scale_colour_manual(
    values = c("red", "green", "blue"),
    labels = c("Chain 1", "Chain 2", "Chain 3")
  ) +
  labs(title = latex2exp::TeX("Trace Plot: $\\sigma_\\epsilon$"),
       x = "m", y = latex2exp::TeX("$\\sigma_\\epsilon$")) +
  p_theme

legend <- gtable::gtable_filter(ggplot_gtable(ggplot_build(p2)), "guide-box")

grid.arrange(
  p1 + theme(legend.position = "none"), 
  p2 + theme(legend.position = "none"), 
  p3 + theme(legend.position = "none"), 
  p4 + theme(legend.position = "none"), 
  p5 + theme(legend.position = "none"), 
  p6 + theme(legend.position = "none"), 
  p7 + theme(legend.position = "none"), 
  p8 + theme(legend.position = "none"),
  p9 + theme(legend.position = "none"),
  p10 + theme(legend.position = "none"),
  ncol = 2, legen = legend,
  heights = c(1.1, 1.1, 1.1, 1.1, 1.1, 0.1)
)
```

```{r echo=FALSE, fig.align='center', fig.cap="Random walk state model: Gaussian kernel density estimates of the marginal posterior distributions", fig.height=8.5, fig.width=6, warning=FALSE, extension-ardl-model-marginal-density, eval=FALSE}
traces <- rbind(
  result[[1]]@traces, 
  result[[2]]@traces, 
  result[[3]]@traces
)

# Plot theme:
p_theme <- theme(
  legend.justification = c(0, 1), legend.position = c(0, 1), 
  legend.title = element_blank(), 
  legend.background = element_rect(fill = "transparent"),
  plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"),
  plot.title = element_text(hjust = 0.5, size = 8),
  axis.title.x = element_text(size = 8),
  axis.title.y = element_text(size = 8, vjust = 0.5)
)

# beta_0
df <- data.frame(time = 1:nrow(traces), beta_0 = traces[, 3])
p1 <- ggplot(data = df, aes(x = beta_0)) +
  geom_density(aes(x = beta_0, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = mean(traces[, 3]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", 
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, "Prior" = 1), name = ""
  ) + 
  labs(title = latex2exp::TeX(
    "Marginal Posterior: \\textbf{$\\beta_0$}"
  ), x = latex2exp::TeX("$\\beta_0$"))

# beta_1
df <- data.frame(time = 1:nrow(traces), beta_1 = traces[, 4])
p2 <- ggplot(data = df, aes(x = beta_1)) +
  geom_density(aes(x = beta_1, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = mean(traces[, 4]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", 
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, "Prior" = 1), name = ""
  ) + 
  labs(title = latex2exp::TeX(
    "Marginal Posterior: \\textbf{$\\beta_1$}"
  ), x = latex2exp::TeX("$\\beta_1$"))

# beta_2
df <- data.frame(time = 1:nrow(traces), beta_2 = traces[, 5])
p3 <- ggplot(data = df, aes(x = beta_2)) +
  geom_density(aes(x = beta_2, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = mean(traces[, 5]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", 
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, "Prior" = 1), name = ""
  ) + 
  labs(title = latex2exp::TeX(
    "Marginal Posterior: \\textbf{$\\beta_2$}"
  ), x = latex2exp::TeX("$\\beta_2$"))

# beta_3
df <- data.frame(time = 1:nrow(traces), beta_3 = traces[, 6])
p4 <- ggplot(data = df, aes(x = beta_3)) +
  geom_density(aes(x = beta_3, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = mean(traces[, 6]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", 
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, "Prior" = 1), name = ""
  ) + 
  labs(title = latex2exp::TeX(
    "Marginal Posterior: \\textbf{$\\beta_3$}"
  ), x = latex2exp::TeX("$\\beta_3$"))

# beta_4
df <- data.frame(time = 1:nrow(traces), beta_4 = traces[, 7])
p5 <- ggplot(data = df, aes(x = beta_4)) +
  geom_density(aes(x = beta_4, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = mean(traces[, 7]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", 
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, "Prior" = 1), name = ""
  ) + 
  labs(title = latex2exp::TeX(
    "Marginal Posterior: \\textbf{$\\beta_4$}"
  ), x = latex2exp::TeX("$\\beta_4$"))

# beta_5
df <- data.frame(time = 1:nrow(traces), beta_5 = traces[, 8])
p6 <- ggplot(data = df, aes(x = beta_5)) +
  geom_density(aes(x = beta_5, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = mean(traces[, 8]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", 
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, "Prior" = 1), name = ""
  ) + 
  labs(title = latex2exp::TeX(
    "Marginal Posterior: \\textbf{$\\beta_5$}"
  ), x = latex2exp::TeX("$\\beta_5$"))

# sigma_u
df <- data.frame(time = 1:nrow(traces), sigma_u = exp(traces[, 9]))
p7 <- ggplot(data = df, aes(x = sigma_u)) +
  geom_density(aes(x = sigma_u, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = mean(exp(traces[, 9])), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", 
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, "Prior" = 1), name = ""
  ) + 
  labs(title = latex2exp::TeX("Marginal Posterior: \\textbf{$\\sigma_u$}"),
       x = latex2exp::TeX("$\\sigma_u$"))

# sigma_epsilon
df <- data.frame(time = 1:nrow(traces), sigma_e = exp(traces[, 10]))
p8 <- ggplot(data = df, aes(x = sigma_e)) +
  geom_density(aes(x = sigma_e, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = mean(exp(traces[, 10])), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", 
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, "Prior" = 1), name = ""
  ) + 
  labs(title = latex2exp::TeX(
    "Marginal Posterior: \\textbf{$\\sigma_\\epsilon$}"
  ), x = latex2exp::TeX("$\\sigma_\\epsilon$"))

# e_lpd_0
df <- data.frame(time = 1:nrow(traces), beta_0 = traces[, 3])
p9 <- ggplot(data = df, aes(x = beta_0)) +
  geom_density(aes(x = beta_0, colour = "Posterior", linetype = "Posterior"), 
               show.legend = FALSE) + 
  geom_vline(
    aes(xintercept = mean(traces[, 3]), 
        colour = "Posterior Mean", linetype = "Posterior Mean"), 
    show.legend = FALSE
  ) +
  stat_function(
    fun = function(...){1}, aes(colour = "Prior", linetype = "Prior")
  ) +
  theme_light() + p_theme + 
  scale_color_manual(
    values = c("Posterior" = "black", "Prior" = "#30914a", 
               "Posterior Mean" = "#2126b5"), 
    name = ""
  ) +
  scale_linetype_manual(
    values = c("Posterior" = 1, "Posterior Mean" = 3, "Prior" = 1), name = ""
  ) + 
  labs(title = latex2exp::TeX(
    "Marginal Posterior: \\textbf{$\\tilde{\\eta}_0$}"
  ), x = latex2exp::TeX("$\\tilde{\\eta}_0$"))

get_legend <- function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}
legend <- get_legend(p2)

grid.arrange(
  p1 + theme(legend.position = "none"), 
  p2 + theme(legend.position = "none"), 
  p3 + theme(legend.position = "none"), 
  p4 + theme(legend.position = "none"), 
  p5 + theme(legend.position = "none"), 
  p6 + theme(legend.position = "none"), 
  p7 + theme(legend.position = "none"), 
  p8 + theme(legend.position = "none"),
  p9 + theme(legend.position = "none"),
  ncol = 2, legend = legend,
  heights = c(1.1, 1.1, 1.1, 1.1, 0.1)
)
```

# Appendix: Algorithms {#algorithms}

The following algorithms can be found in [@pomp_article].

## Particle filter and systematic resampling

\begin{algorithm}
  \caption{Particle Filter}
  \label{alg:particle-filter}
    \KwIn {Simulators $\mu_{\theta}(x_{0})$ and $f_{\theta}(x_{n}|x_{n-1})$ \newline
    Evaluator $g_{\theta}(y_{n}|x_{n})$ \newline
    Parameter $\theta$ \newline
    Data $y^*_{1:N}$ \newline
    Number of particles $J$
    }
  \nl Initialize filter particles: simulate $X^F_{0,j} \sim \mu_{\theta}(\cdot)$ for $j$ in $1:J$\;
  \nl \For{$n = 1$ \To{} $N$}{%
    \nl Simulate for prediction: $X^P_{n,j} \sim f_{\theta}(\cdot|X^F_{n-1,j})$ for $j$ in $1:J$\;
    \nl Evaluate weights: $w(n, j) = g_{\theta}(y^*_{n}|X^P_{n,j})$ for $j$ in $1:J$\;
    \nl Normalize weights: $\tilde{w}(n, j) = \frac{w(n, j)}{\sum^J_{m=1}w(n, m)}$\;
    \nl Apply Algorithm \ref{alg:systematic-resampling} to select indices $k_{1:J}$ with $P(k_j = m) = \tilde{w}(n, m)$\;
    \nl Resample: set $X^F_{n,j} = X^P_{n,k_j}$ for $j$ in $1:J$\;
    \nl Compute conditional log likelihood: $\hat{\ell}_{n|n-1} = ln\Big(\frac{1}{J}\sum^J_{m=1}w(n, m)\Big)$\;
  }
  \KwOut{Log likelihood estimate: $\hat{\ell}(\theta) = \sum_{n=1}^N\hat{\ell}_{n|n-1}$ \newline
  Filter sample: $X^F_{n,1:J}$ for $n$ in 1:N}
\end{algorithm}

\begin{algorithm}
  \caption{Systematic Resampling}
  \label{alg:systematic-resampling}
    \KwIn {Weights $\tilde{w}_{1:J}$, normalized so that $\sum^J_{j=1}\tilde{w}_{j} = 1$
    }
  \nl Construct cumulative sum: $c_j = \sum_{m=1}^j \tilde{w}_m$ for $j$ in $1:J$\;
  \nl Draw a uniform initial sampling point: $U_1 \sim \mathcal{U}(0, \frac{1}{J})$\;
  \nl Construct evenly spaced sampling points: $U_j = U_1 + (j - 1)\frac{1}{J}$ for $j$ in $2:J$\;
  \nl Initialize: set $p=1$\;
  \nl \For{$j = 1$ \To{} $J$}{%
    \nl \While{$U_j > c_p$}{%
      \nl Step to the next resampling index: set $p = p + 1$\;
    }
    \nl Assign resampling index: set $k_j = p$\;
  }
  \KwOut{Resampling indices $k_{1:J}$}
\end{algorithm}

## PMMH

\begin{algorithm}
  \caption{PMMH}
  \label{alg:pmmh}
    \KwIn {Simulators $\mu_{\theta}(x_{0})$ and $f_{\theta}(x_{n}|x_{n-1})$ \newline
    Evaluator $g_{\theta}(y_{n}|x_{n})$ \newline
    Proposal distribution $q(\theta^*|\theta^{(m-1)})$ \newline
    Starting Parameter $\theta_0$ \newline
    Prior $\pi(\theta)$ \newline
    Data $y^*_{1:N}$ \newline
    Number of particles $J$ \newline
    Number of iterations $M$
    }
  \nl Initialization: Compute $\hat{\ell}_0 = \hat{\ell}(\theta_0)$ using Algorithm \ref{alg:particle-filter} with $J$ particles\;
  \nl Draw random trajectory $X^{(0)}_{0:N}$from $X^F_{0:N,1:J}$ obtained from particle filter \;
  \nl \For{$m = 1$ \To{} $M$}{%
    \nl Draw a parameter proposal $\theta^*\sim q(\theta^*|\theta^{(m-1)})$\;
    \nl Compute $\hat{\ell}^* = \hat{\ell}(\theta^*)$ using algorithm \ref{alg:particle-filter} with $J$ particles\;
    \nl Draw random trajectory $X^{*}_{0:N}$ from $X^F_{0:N,1:J}$ obtained from particle filter \;
    \nl Compute acceptance probability $$\alpha(\theta^{*}|\theta^{(m-1)}) = \min\Bigg\{1, \frac{\text{exp}(\hat{\ell}^*)(\theta^*)\pi(\theta^*)q(\theta^{(m-1)}|\theta^*)}{\text{exp}(\hat{\ell}^{(m-1)})\pi(\theta^{(m-1)})q(\theta^*|\theta^{(m-1)})}\Bigg\}$$\;
    \nl Generate $U \sim \mathcal{U}(0, 1)$\;
    \nl Set $$
    (\theta^{(m)}, \hat{\ell}^{(m)}, X^{(m)}_{0:N}) =
    \begin{cases}
      (\theta^{*}, \hat{\ell}^{*}, X^{*}_{0:N}), & \text{if}\;\; U < \alpha(\theta^{*}|\theta^{(m-1)}) \\
      (\theta^{(m-1)}, \hat{\ell}^{(m-1)}, X^{(m-1)}_{0:N}), & \text{otherwise}
    \end{cases}$$\;
  }
  \KwOut{Samples $X^{(1:M)}_{0:N}$, $\theta^{(1:M)}$ from posterior distribution $p(x_{0:N}, \theta|y^*_{1:N})$}
\end{algorithm}

# Appendix: Derivations {#derivations}

## Prediction & Filtering distribution

\begin{equation*}
\begin{split}
  p_{\theta}(x_n|y^*_{1:n-1}) & = \int p_{\theta}(x_n, x_{n-1}|y^*_{1:n-1})dx_{n-1} \\
  & = \int \frac{p_{\theta}(x_n, x_{n-1}, y^*_{1:n-1})}{p_{\theta}(y^*_{1:n-1})}dx_{n-1} \\
  & = \int \frac{p_{\theta}(x_n, y^*_{1:n-1} | x_{n-1})p_{\theta}(x_{n-1})}{p_{\theta}(y^*_{1:n-1})}dx_{n-1} \\
  & = \int \frac{p_{\theta}(x_n| y^*_{1:n-1}, x_{n-1})p_{\theta}(y^*_{1:n-1}, x_{n-1})p_{\theta}(x_{n-1})}{p_{\theta}(y^*_{1:n-1})p_{\theta}(x_{n-1})}dx_{n-1} \\
  & = \int f_{\theta}(x_n|x_{n-1})p_{\theta}(x_{n-1}|y^*_{1:n-1})dx_{n-1}
\end{split}
\end{equation*}

\begin{equation*}
\begin{split}
  p_{\theta}(x_n|y^*_{1:n}) & = p_{\theta}(x_n|y^*_n, y^*_{1:n-1}) = \frac{
    p_{\theta}(y^*_n, y^*_{1:n-1} | x_{n})p_{\theta}(x_n)
  }{
    p_{\theta}(y^*_n, y^*_{1:n-1})
  } \\
  & = \frac{
    p_{\theta}(y^*_n | y^*_{1:n-1}, x_{n})p_{\theta}(y^*_{1:n-1} | x_n)p_{\theta}(x_n)
  }{
    p_{\theta}(y^*_n | y^*_{1:n-1})p_{\theta}(y^*_{1:n-1})
  } \\
  & = \frac{
    p_{\theta}(y^*_n | y^*_{1:n-1}, x_{n})p_{\theta}(x_n | y^*_{1:n-1})p_{\theta}(y^*_{1:n-1})p_{\theta}(x_n)
  }{
    p_{\theta}(y^*_n|y^*_{1:n-1})p_{\theta}(y^*_{1:n-1})p_{\theta}(x_n)
  } \\
  & = \frac{g_{\theta}(y^*_n|x_{n})p_{\theta}(x_n|y^*_{1:n-1})}{p_{\theta}(y^*_n|y^*_{1:n-1})}
\end{split}
\end{equation*}

## PMMH - acceptance probability

\begin{equation*}
\begin{split}
\alpha(\theta^{*}|\theta^{(m-1)}) & = \min\Bigg\{1, \frac{p(\theta^*, x^*_{0:N}|y^*_{1:N})q(\theta^{(m-1)}, x^{(m-1)}_{0:N}|\theta^*, x^*_{0:N})}{p(\theta^{(m-1)}, x^{(m-1)}_{0:N}|y^*_{1:N})q(\theta^*, x^*_{0:N}|\theta^{(m-1)}, x^{(m-1)}_{0:N})}\Bigg\} \\
& = \min\Bigg\{1, \frac{p(\theta^*, x^*_{0:N}|y^*_{1:N})q(\theta^{(m-1)}|\theta^*)p_{\theta^{(m-1)}}(x^{(m-1)}_{0:N}|y^*_{1:N})}{p(\theta^{(m-1)}, x^{(m-1)}_{0:N}|y^*_{1:N})q(\theta^*|\theta^{(m-1)})p_{\theta^*}(x^*_{0:N}|y^*_{1:N})}\Bigg\} \\
& = \min\Bigg\{1, \frac{\mathcal{L}(\theta^*)\pi(\theta^*)q(\theta^{(m-1)}|\theta^*)}{\mathcal{L}(\theta^{(m-1)})\pi(\theta^{(m-1)})q(\theta^*|\theta^{(m-1)})}\Bigg\}
\end{split}
\end{equation*}

Using that:

\begin{equation*}
\begin{split}
p_{\theta}(y_{1:N}) & = \frac{p_{\theta}(y_{1:N}|x_{0:N})p_{\theta}(x_{0:N})}{p_{\theta}(x_{0:N}|y_{1:N})} \\
& = \frac{p(y_{1:N}, x_{0:N}, \theta)p(x_{0:N}, \theta)p(y_{1:N}, \theta)}{p(y_{1:N}, x_{0:N}, \theta)p(x_{0:N}, \theta)p(\theta)} \\
& = \frac{p(y_{1:N}, \theta)}{p(\theta)}
\end{split}
\end{equation*}

\begin{equation*}
p(\theta, x_{0:N}|y_{1:N}) \propto p_{\theta}(y_{1:N}|x_{0:N})p_{\theta}(x_{0:N})\pi(\theta)
\end{equation*}

